---
title: "My R makrdown Project"
author: "Kailash Natarajan"
date: "19/06/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading the data
```{r}
datafile<-read.csv("activity.csv")
```
What is mean total number of steps taken per day?
```{r}
##Ignoring missing values and putting into a new variable
temp<-datafile
temp[is.na(datafile[,1]),1]<-0
library(ggplot2)
step_sum<-tapply(temp$steps,temp$date,sum)
##1. The total number of steps taken per day
print("Steps taken per day")
print(step_sum)
##2. Histogram of number of steps
hist(step_sum)
##3. Finding mean and median values
##Finding mean and median of steps taken per day
#This will return only the values
values<-as.numeric(step_sum)
print("Mean is:")
print(mean(values))
print("Median is:")
print(median(values))
```
What is the average daily activity pattern?
```{r}
##Ignoring removing NA and making it 0
interval_mean<-tapply(temp$steps,temp$interval,mean)
plot(interval_mean~names(interval_mean),type="l",xlab="Interval",ylab="Average number of steps")
print("Maximum number of average steps in interval:")
print(interval_mean)
mx<-max(interval_mean)
index<-which(mx==interval_mean)
print(names(interval_mean)[index[[1]]])
```
Imputing missing values
```{r}
#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA's)
nas<-sum(is.na(datafile$steps)|is.na(datafile$interval))
print("Number of rows with NA:")
print(nas)
#Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
##Finding mean of values in the same interval and using it to
##impute the NA's
new_datafile<-datafile
int_mean<-tapply(temp$steps,temp$interval,mean)
for(i in unique(datafile$interval)){
  new_datafile[is.na(new_datafile$steps)&(new_datafile$interval==i),"steps"]<-as.numeric(int_mean[names(int_mean)==i])
}
##Histogram, mean and median of new datafile
step_sum<-tapply(new_datafile$steps,new_datafile$date,sum)
hist(step_sum)
print("Mean:")
values<-as.numeric(step_sum)
print(mean(values))
print("Median:")
print(median(values))

```
From the above we can see that mean has changed slightly and median has not. This implies that by using even a simple model for imputing like just converting NA's to respective means, we can solve the issue of missing values without impacting the calculations and analysis too much.

Are there differences in activity patterns between weekdays and weekends?
```{r}
library(dplyr)
##Converting date field to date class
new_datafile$date<-as.Date(as.character(new_datafile$date),"%Y-%m-%d")
new_datafile<-mutate(new_datafile,week=weekdays(date))
new_datafile$week<-sapply(new_datafile$week,                          function(x){if(x=="Saturday"|x=="Sunday"){y<-"Weekend"}
  else{y<-"Weekday"}
  y
  })
activity_by_date <- aggregate(steps~interval + week, new_datafile, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = week)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~week, ncol = 1, nrow=2)
print(plot)
```